---
title: "YWAR Structure"
author: "Rachael Bay"
date: "5/3/2017"
output: html_document
---

```{r setup, include=FALSE,warnings=FALSE,message=F}
library(RColorBrewer)
library(rgdal)
library(tess3r) #This is Eric's version of Tess for mapping rasters
library(tidyverse)
library(raster)
library(ggmap)
library(ggspatial)
library(knitr)
knitr::opts_chunk$set(dev="png")

####Read in qmatrices
K2 <- read.delim("../structure_out/96snps_qmat/randomrun/K2.txt",row.names=1,header=F)
K3 <- read.delim("../structure_out/96snps_qmat/randomrun/K3.txt",row.names=1,header=F)
K4 <- read.delim("../structure_out/96snps_qmat/randomrun/K4.txt",row.names=1,header=F)
K5 <- read.delim("../structure_out/96snps_qmat/randomrun/K5.txt",row.names=1,header=F)

###Read and order meta
raw <- read.delim("../data/metadata/Fluidigm_meta_03.20.17.txt")
meta <- raw[match(rownames(K2),raw$Field_Number),]
dim(meta)
meta$Pop[is.na(meta$Pop)] <- max(meta$Pop,na.rm=T)+1  ##I had 5 samples from NJ without a Pop number

###Remove the one Thunder Bay sample - I think it's a migrant!
K2 <- K2[meta$Near_Town!="Thunder Bay",]
K3 <- K3[meta$Near_Town!="Thunder Bay",]
K4 <- K4[meta$Near_Town!="Thunder Bay",]
K5 <- K5[meta$Near_Town!="Thunder Bay",]
meta <- meta[meta$Near_Town!="Thunder Bay",]

```

##Set up the data

We'll start with q-matrix output from structure for a range of K values formatted like this:

```{r qmatrix,warnings=FALSE}
head(K2)
```

We also have metadata, in the same order:

```{r meta}
head(meta)
```

For now, the `Pop` column is arbitrary, but now we'll create new population numbers based on longitude - this will be helpful for ordering samples in the barplot

```{r fakepop,warnings=FALSE,message=F}
popcoords <- aggregate(meta[,c("Long","Lat")],list(meta$Pop),mean)
ordercoords <- popcoords[order(popcoords$Long),]
ordercoords$Fakepop <- 1:nrow(ordercoords)
meta$Fakepop <-ordercoords$Fake[match(meta$Pop,ordercoords$Group.1)]
```

## Barplots

```{r barparams, include=FALSE,warnings=FALSE,message=F}
###These loops calculate breakpoints between populations
sizes <- table(meta$Fakepop)#*1.200704 ###Need to figure out scaling factor manually
splits <- c(0)
for (i in 1:length(sizes)) {
  new <- splits[i]+sizes[i]
  splits <- c(splits,new)
}
mids <- c()
for (i in 1:(length(sizes))) {
  half <- (splits[i]+splits[i+1])/2
  mids <- c(mids,half)
}

###Plotting parameters for barplots
par(mar=c(1,1,1,1),oma=c(10,2,1,1))
colors <- brewer.pal(8,"Set2")
```

We can reorder the samples by our new pop and plot:

```{r barplot1,warnings=FALSE,message=F}
K2order <- K2[order(meta$Fakepop),]
barplot(t(K2order),col=c("red","blue"))
```

Let's plot all the K together and make it pretty!
```{r barplot2,echo=F,warnings=FALSE,message=F}
K2order <- K2[order(meta$Fakepop),]
K3order <- K3[order(meta$Fakepop),]
K4order <- K4[order(meta$Fakepop),]
K5order <- K5[order(meta$Fakepop),]
ordermeta <- meta[order(meta$Fakepop),]

par(mfrow=c(4,1),mar=c(1,1,1,1),oma=c(10,2,1,1))
barplot(t(K2order),col=colors,border=NA,axes=F,names.arg=rep("",nrow(K2)),space=0)
abline(v=splits,lwd=0.75)
barplot(t(K3order),col=colors,border=NA,axes=F,names.arg=rep("",nrow(K2)),space=0)
abline(v=splits,lwd=0.75)
barplot(t(K4order),col=colors,border=NA,axes=F,names.arg=rep("",nrow(K2)),space=0)
abline(v=splits,lwd=0.75)
barplot(t(K5order),col=colors,border=NA,axes=F,names.arg=rep("",nrow(K2)),space=0)
abline(v=splits,lwd=0.75)
axis(1,at=mids,labels=ordercoords$Fakepop,las=2,cex.axis=1.3,tck=0,lwd=0)
```

##Maps

Now we want to spatially interpolate structure results onto a map

First we read in the shape file
```{r shapefile,warnings=FALSE,message=F,dpi=36, out.width="600px", out.height="600px"}
range <- readOGR("../data/shapefile/",layer="YWAR")
breeding <- subset(range,SEASONAL==2)
breeding <- crop(breeding,extent(-170,-55,30,70))
palette <- CreatePalette(brewer.pal(8,"Set2"))

plotK <- K3order[,1:3]
stack <- tess3Q_map_rasters(as.matrix(plotK),as.matrix(ordermeta[,c("Long","Lat")]),method="map.max", interpol = FieldsKrigModel(10),  
     main = "Ancestry coefficients",
     xlab = "Longitude", ylab = "Latitude", 
     resolution = c(300,300), cex = .4,
     col.palette = palette, map.polygon = breeding)
#plot(stack)
names(stack) <- paste0("grp",1:ncol(plotK))
maxes <- max(stack)
keep <- stack>=maxes
stack[keep<=0] <- NA

long_rast <- lapply(names(stack), function(n) {
  r <- stack[[n]]
  xy <- as_tibble(xyFromCell(r, 1:ncell(r)))
  v <- as.data.frame(r)
  bind_cols(xy, v) %>%
    setNames(c("long", "lat", "value")) %>%
    mutate(grp = n)
}) %>%
  bind_rows() %>%
  filter(!is.na(value))

long_rast2 <- long_rast %>%
  mutate(value = ifelse(value > 1, 1.0, value)) %>%   # force everything to be no more than 1
  mutate(value = value ^2)


# Load some map layers and things
nat_earth <- stack("../data/maps/HYP_LR_SR_W_DR.tif")
ne_coast <- readOGR("../data/maps/ne_10m_coastline",
                    "ne_10m_coastline")
state_prov <- readOGR("../data/maps/ne_10m_admin_1_states_provinces_lines",
                      "ne_10m_admin_1_states_provinces_lines")

# and immediately drop all but canada and the US:
state_prov <- state_prov[state_prov$adm0_name %in% c("United States of America", "Canada"),]
country_bound <- readOGR("../data/maps/ne_10m_admin_0_boundary_lines_land",
                         "ne_10m_admin_0_boundary_lines_land")

###Subsetting
# A function for subsetting spatial objects and making them long format
# here eric implements the same thing but using tidyverse tools
tidy_subset <- function(x, longlat) {
  x@data$id <- rownames(x@data)
  x.f <- broom::tidy(x) %>%
    dplyr::left_join(., x@data, by = "id") %>%
    dplyr::tbl_df() %>%
    filter(long > longlat[1],
           long < longlat[2],
           lat > longlat[3],
           lat < longlat[4])
}

#### Now Clip some things to the desired range, and longify the nat_earth raster ####
domain <- c(-170, -55, 30, 70)
coast_subset <- tidy_subset(ne_coast, domain)
state_subset <- tidy_subset(state_prov, domain)
country_subset <- tidy_subset(country_bound, domain)
nat_crop <- crop(nat_earth, y = extent(domain))
rast_table <- data.frame(xyFromCell(nat_crop, 1:ncell(nat_crop)),
                         getValues(nat_crop/255))
rast_table$rgb <- with(rast_table, rgb(HYP_LR_SR_W_DR.1,
                                       HYP_LR_SR_W_DR.2,
                                       HYP_LR_SR_W_DR.3,
                                       1))

#### Make the base map  ####
base_map <- ggplot() +
  geom_raster(data = rast_table, mapping = aes(x = x, y = y), fill = rast_table$rgb, interpolate = TRUE) +
  geom_path(data=state_subset, aes(x = long, y = lat, group = group), color = 'gray30', size = 0.2) +
  geom_path(data=country_subset, aes(x = long, y = lat, group = group), color = 'gray30', size = 0.4) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_quickmap(xlim = domain[1:2],  ylim = domain[3:4]) +
  ylab("Latitude") +
  xlab("Longitude") +
  theme(axis.text.x = element_text(size=15),
        axis.text.y = element_text(size=15))

structmap <- base_map + 
  geom_raster(data = long_rast2, aes(x = long, y = lat, fill = grp, alpha = value), na.rm = TRUE) +
  scale_colour_manual(values = c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628')) +
  scale_alpha_continuous(range = c(0.0, 1))

###Add sampling points
ordercoords$n <- sapply(ordercoords[,1],function(x) length(which(meta$Fakepop==x)))
plotcoords <- ordercoords[ordercoords$n>0,]
plotcoords$adjLong <- plotcoords$Long+2
plotcoords$adjLat <- plotcoords$Lat+1
addpoints <- structmap + 
  geom_point(data=plotcoords,mapping=aes(x=Long,y=Lat),colour="white",alpha=0.8) +
  geom_text(data=plotcoords,mapping=aes(x=adjLong,y=adjLat,label=Group.1),size=3)
plot(addpoints)

```
